Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-19 WSSSPE2_report/WSSSPE2_review.tex
minor fix and adding lightning talk for myself

%\item Daniel S. Katz: Implementing Transitive Credit with JSON-LD

@@

After this discussion, group A brainstormed about incentives,
with the following items suggested:
\begin{itemize}
\item running programming contests, creating bounties for contributing to open
source software, etc.
\item augment author lists to give credit to people who don't now get credit
(and making them machine readable)
\item developing a microcitation standard and mechanism (for both software and
data)
\item developing a well-defined standard for author contribution -- what level
of contribution rises to the level of authorship?
\item leveraging social media for citation and reviewing of content -- then
using social media to bring more people into the review process than is
traditional
\item determining where else software can be cited and recorded (e.g.,
acknowledgments sections of papers)
\item developing a taxonomy of contributors (e.g., Project
Credit~\cite{projectcredit}) tied to places that these metrics are already
stored (e.g., ORCID) \cite{orcid}
\item making metadata easier to add for software, creating incentive for
providing software metadata -- note that this cannot be centralized
\item creating something like the h-index that tenure committees can make use of
- simplify a way of measuring and documenting the overall credit given to an
individual over different projects
\item thinking about publishing software versus journal - software does not have to
be novel
\item determining guidelines for recommending software characteristics for
tenure - perhaps draft guidelines then get ACM or IEEE agreement
\end{itemize}

@@

The discussion of incentivizing those in science acknowledged that the
publications system was far from perfect for incentivizing good software work.
Nonetheless discussion focused on bringing software people into publications.
There were two main suggestions. The first is to focus on end users of software
and encourage them to cite software the ``right" way. James Howison suggested
that his research showed that few projects were making a formal request for
citation (but that authors weren't necessarily following those suggestions
anyway) \cite{howison2015jasist}. He suggested making access to the software
conditional on a license that requires citation. Others found this ``too
confrontational" and preferred to concentrate on making it easier to do the
right thing. The second was focusing on those leading software projects, and the
group was more enthusiastic about ``forcing" PIs to include their ``software
people" on publications, although there were few ideas on how exactly to do
this. Another technique mentioned was that when scientific software projects are
hosted in organizations like Apache, the scientific contributors can benefit
from building their reputations, perhaps yielding job offers that they can use
to negotiate better job and career packages.

>>>>

\item \textbf{Daniel S. Katz: Implementing Transitive Credit with JSON-LD.}
Science and engineering research increasingly relies on activities that facilitate research but are not currently rewarded or recognized, such as: data sharing; developing common data resources, software and methodologies; and annotating data and publications. To promote and advance these activities, we must develop mechanisms for assigning credit, facilitate the appropriate attribution of research outcomes, devise incentives for activities that facilitate research, and allocate funds to maximize return on investment. Katz discussed the issue of assigning credit for both direct and indirect contributions by using JSON-LD to implement a prototype transitive credit system~\cite{Katz_transitive_credit_poster}.

@@

After this discussion, group A brainstormed about incentives,
with the following items suggested:
\begin{itemize}
\item running programming contests, creating bounties for contributing to open
source software, etc.
\item augment author lists to give credit to people who don't now get credit
(and making them machine readable)
\item developing a microcitation standard and mechanism (for both software and
data)
\item developing a well-defined standard for author contribution -- what level
of contribution rises to the level of authorship?
\item leveraging social media for citation and reviewing of content -- then
using social media to bring more people into the review process than is
traditional
\item determining where else software can be cited and recorded (e.g.,
acknowledgments sections of papers)
\item developing a taxonomy of contributors (e.g., Project
Credit~\cite{projectcredit}) tied to places that these metrics are already
stored (e.g., ORCID~\cite{orcid})
\item making metadata easier to add for software, creating incentive for
providing software metadata -- note that this cannot be centralized
\item creating something like the h-index that tenure committees can make use of
- simplify a way of measuring and documenting the overall credit given to an
individual over different projects
\item thinking about publishing software versus journal - software does not have to
be novel
\item determining guidelines for recommending software characteristics for
tenure - perhaps draft guidelines then get ACM or IEEE agreement
\end{itemize}

@@

The discussion of incentivizing those in science acknowledged that the
publications system was far from perfect for incentivizing good software work.
Nonetheless discussion focused on bringing software people into publications.
There were two main suggestions. The first is to focus on end users of software
and encourage them to cite software the ``right" way. James Howison suggested
that his research showed that few projects were making a formal request for
citation (but that authors weren't necessarily following those suggestions
anyway)~\cite{howison2015jasist}. He suggested making access to the software
conditional on a license that requires citation. Others found this ``too
confrontational" and preferred to concentrate on making it easier to do the
right thing. The second was focusing on those leading software projects, and the
group was more enthusiastic about ``forcing" PIs to include their ``software
people" on publications, although there were few ideas on how exactly to do
this. Another technique mentioned was that when scientific software projects are
hosted in organizations like Apache, the scientific contributors can benefit
from building their reputations, perhaps yielding job offers that they can use
to negotiate better job and career packages.
